\chapter{Extended Methods}
\label{appendix:extended-methods}

\section{Flies}

Flies that expressed CsChrimson were raised on cornmeal--molasses food (Archon Scientific) supplemented with 100~\textmu l all-trans-retinal (Sigma; 17~mM in ethanol) and kept wrapped in light-blocking foil to avoid photoconversion. Flies that did not express CsChrimson were raised on standard cornmeal--molasses food. All flies were housed in an incubator under a 12~h:12~h light:dark cycle at 25~\textdegree C and 50--70\% relative humidity.

Experimenters were aware of fly genotype in all cases except for the hDeltaB optogenetic stimulation experiment. \textbf{TODO: add reference to the corresponding main-text figure.} Sample sizes were selected according to field-standard conventions, which are typically based on the anticipated degree of animal-to-animal variability as informed by published studies and pilot experiments. All experiments were performed using flies carrying at least one wild-type copy of the white (\textit{w}) gene.

\paragraph{Genotypes used in each figure.}

{\raggedright
\begin{itemize}
    \item MBON21 calcium imaging: \textit{w}/+; +; \textit{P}\{y[+t7.7] w[+mC]=VT999036-GAL4\}attP2/\textit{PBac}\{y[+t7.7] \textit{PBac}\{y[+mDint2] w[+mC]=20XUAS-IVS-jGCaMP7f\}VK00005.
    \item MBON09 calcium imaging: \textit{w}/+; +; \textit{P}\{y[+t7.7] w[+mC]=R94B10-GAL4.DBD\}attP2 \textit{PBac}\{y[+mDint2] w[+mC]=R20A02-p65.AD\}VK00027/\textit{PBac}\{y[+t7.7] \textit{PBac}\{y[+mDint2] w[+mC]=20XUAS-IVS-jGCaMP7f\}VK00005.
    \item hDeltaB calcium imaging: \textit{w}/+; +/\textit{P}\{R72B05-p65.AD\}; attP40\textit{P}\{y[+t7.7] w[+mC]=UAS-sytjGCaMP7f\}attP2;/\textit{P}\{VT055827-Gal4.DBD\}attP2.
    \item hDeltaB and MBON09 calcium imaging: \textit{w}/+; \textit{P}\{y[+t7.7] w[+mC]=UAS-Syt1::jGCaMP7f\}attP40/\textit{P}\{R72B05-p65.AD\}attP40; \textit{P}\{y[+t7.7] w[+mC]=GMR94B10-GAL4\}attP2/\textit{P}\{VT055827-Gal4.DBD\}attP2.
    \item hDeltaB and MBON21 calcium imaging: \textit{w}/+; \textit{P}\{y[+t7.7] w[+mC]=UAS-Syt1::jGCaMP7f\}attP40/\textit{P}\{R72B05-p65.AD\}attP40; \textit{P}\{y[+t7.7] w[+mC]=VT999036-GAL4\}attP2/\textit{P}\{VT055827-Gal4.DBD\}attP2.
    \item hDeltaB optogenetics: \textit{w}/+; \textit{P}\{y[+t7.7] w[+mC]=20XUAS-IVS-CsChrimson.mVenus\}attP40/\textit{P}\{R72B05-p65.AD\}attP40; \textit{P}\{VT055827-Gal4.DBD\}attP2/+.
    \item MBON21 optogenetics and imaging: \textit{w}/+; \textit{PBac}\{y[+mDint2] w[+mC]=20X-UAS-IVS-CsChrimson::mCyRFP3\}VK00002/+; \textit{P}\{y[+t7.7] w[+mC]=VT999036-GAL4\}attP2/\textit{PBac}\{y[+mDint2] w[+mC]=20XUAS-IVS-Syt1::jGCaMP8m\}VK00005.
    \item MBON09 optogenetics and imaging: \textit{w}/+; \textit{PBac}\{y[+mDint2] w[+mC]=20X-UAS-IVS-CsChrimson::mCyRFP3\}VK00002/+; \textit{P}\{y[+t7.7] w[+mC]=R94B10-GAL4.DBD\}attP2 \textit{PBac}\{y[+mDint2] w[+mC]=R20A02-p65.AD\}VK00027/\textit{PBac}\{y[+mDint2] w[+mC]=20XUAS-IVS-Syt1::jGCaMP8m\}VK00005.
\end{itemize}
}

\paragraph{Origins of transgenic stocks.}

{\raggedright
The following stocks were obtained from the Bloomington Drosophila Stock Center (BDSC) and previously published as follows: \textit{P}\{y[+t7.7] w[+mC]=R94B10-GAL4.DBD\}attP2 \textit{PBac}\{y[+mDint2] w[+mC]=R20A02-p65.AD\}VK00027 (BDSC\_68262), \textit{P}\{y[+t7.7] w[+mC]=20XUAS-IVS-CsChrimson.mVenus\}attP40 (BDSC\_55135), \textit{P}\{y[+t7.7] w[+mC]=UAS-sytjGCaMP7f\}attP2 (BDSC\_94619), \textit{P}\{y[+t7.7] w[+mC]=UAS-sytjGCaMP7f\}attP40 (BDSC\_94620), \textit{PBac}\{y[+mDint2] w[+mC]=20XUAS-IVS-CsChrimson::mCyRFP3\}VK00002 (BDSC\_606323).
The following stock was obtained from Yoshinori Aso at Janelia Research Campus: \textit{P}\{y[+t7.7] w[+mC]=VT999036-GAL4\}attP2.
The following stock was obtained from Gaby Maimon at Rockefeller University: \textit{P}\{R72B05-p65.AD\}attP40; \textit{P}\{VT055827-Gal4.DBD\}attP2.
The following stock was obtained from Thomas R.\ Clandinin at Stanford University: \textit{PBac}\{y[+mDint2] w[+mC]=20XUAS-IVS-Syt1::jGCaMP8m\}VK00005.

}

\section{Fly preparation and dissection}

Flies were harvested at least 24 hours before each experiment. For all calcium imaging experiments, they were maintained on standard molasses food, whereas for optogenetic manipulation experiments they were kept on molasses food supplemented with 100~\textmu l all-trans-retinal (Sigma; 17~mM in ethanol). In both types of experiments, only female flies 20--72~h after eclosion were used. For odor-related experiments, flies were food-deprived starting the day before testing and housed with a kimwipe moistened with deionized water to maintain humidity. Experiments were conducted without any circadian timing constraints.

The manual dissection procedure was as follows. Flies were briefly anesthetized on ice and placed with fine forceps (Fine Science Tools) into a custom holder machined from black Delrin (Autotiv or Protolabs). This holder had an inverted pyramid shape to reduce obstruction of the eyes. The head was angled slightly forward to make the posterior side easier to access with the microscope objective. After clipping the wings, the head and thorax were fixed to the holder using UV-curable glue (Aquaseal UV Field Repair Adhesive), which was hardened with a short pulse of ultraviolet light (LED-200, Electro-Lite Co.). To limit large brain movements, the proboscis was also immobilized with a small amount of the same glue. A 30G\,$\times$\,1/2'' needle (BD PrecisionGlide Needle, 30G\,$\times$\,1/2'' Regular Wall Hypodermic, 305106) was then used in extracellular \textit{Drosophila} saline to cut an opening in the head capsule, and tracheoles and fat were removed to expose the brain. Brain motion was further minimized by stretching muscle 16 through gentle pulling on the oesophagus, or alternatively by removing the muscle via an anterior clip. The extracellular saline contained 103~mM NaCl, 3~mM KCl, 5~mM TES, 8~mM trehalose, 10~mM glucose, 26~mM NaHCO$_3$, 1~mM NaH$_2$PO$_4$, 1.5~mM CaCl$_2$, and 4~mM MgCl$_2$, with an osmolarity of 270--275~mOsm. The saline was oxygenated with carbogen (95\%~O$_2$, 5\%~CO$_2$), resulting in a final pH of approximately 7.3.

\section{Visual panorama and visual stimuli}

All calcium imaging and LED-based optogenetics experiments were performed using a custom-built LED display system adapted from the G4 display platform developed by the M.~Reiser lab. Visual stimuli were presented on a circular panoramic arena composed of modular 16\,$\times$\,16 pixel LED panels, with each pixel subtending approximately 1.875\textdegree{} of the fly's visual field. The arena spanned 60\textdegree{} in elevation and 330\textdegree{} in azimuth, with the 30\textdegree{} sector directly behind the animal removed to allow space for ball illumination and tracking hardware. The panels used blue LEDs with a peak emission at 470~nm, selected to minimize spectral overlap with GCaMP fluorescence. For calcium imaging experiments, five layers of Rosco R381 gel filters were placed in front of the arena to further reduce spectral interference, and an additional diffuser layer was added to suppress reflections (SXF-0600, Snow White Light Diffuser, Decorative Films). The visual stimulus consisted of a bright vertical bar with positive contrast, 8 pixels wide (15\textdegree{} azimuth). In most of the arena, the bar extended across the full two-panel height, except in the region spanning $-165$\textdegree{} to $+165$\textdegree{} behind the fly, where only a single display panel was present and the bar height was reduced by half.

In closed-loop experiments, the azimuthal position of the cue was controlled by the fly's yaw rotation measured from the spherical treadmill (see ``Spherical treadmill and locomotion measurement''). A yaw gain of 0.7 was used in all experiments, such that the displacement of the visual cue corresponded to 0.7 times the ball's yaw displacement. In calcium imaging experiments, the cue was pseudorandomly displaced by $\pm 90$\textdegree{} at intervals ranging from 60 to 80~s. After each perturbation, the cue remained stationary for 2~s before closed-loop coupling resumed. Cue position was recorded throughout the experiment using analog output signals from the display panels, which were acquired together with other experimental signals at 20~kHz via a National Instruments PCIe-6363 data acquisition card. During offline analysis, these analog signals were converted into cue position in pixel coordinates.

All two-photon optogenetics experiments were carried out using a custom visual stimulation system implemented in ViRMEn (Virtual Reality MATLAB Engine). Visual stimuli were delivered with a DLP projector (Texas Instruments DLP3010EVM-LC) operating at 120~Hz. To minimize contamination of imaging signals, the projector's red and green LEDs were disabled, and optical band-pass filters (Semrock 480/17-25 and 482/18-25) were placed at the projector output. The projected image was reflected by a 45\textdegree{} first-surface mirror onto a panoramic cone-shaped screen, which covered 60\textdegree{} in elevation and 330\textdegree{} in azimuth, again with the 30\textdegree{} region directly behind the fly omitted to accommodate treadmill illumination and tracking hardware. Stimuli were geometrically corrected to match the curvature of the screen.

In these experiments, the visual stimulus was a bright vertical bar spanning 15\textdegree{} in azimuth on a dark background. Except during heading perturbation trials, this bar served as a celestial cue and was coupled in closed loop exclusively to the fly's rotational movements. The cue position was updated using a gain of 0.7 relative to the fly's movement, consistent with prior work showing that this value improves heading tracking. During heading perturbations, the bar was randomly shifted by 90\textdegree{} to the left or right of its previous position and held at the new location for 2~s before closed-loop control was reinstated. All visual stimuli were controlled through ViRMEn.

\section{Experimental trial structure}

Before data collection, flies were acclimated for at least 15~min in closed loop with the visual cue. For virtual plume-tracking experiments, behavior was recorded in 16--18~min sessions. During each session, the fly remained in closed loop with the cue, which was displaced to a new angular position relative to its current location at random intervals of 60--80~s, alternating between $+90$\textdegree{} and $-90$\textdegree{}.

Two related stimulus designs were used, termed Protocol A and Protocol B. In both protocols, odor delivery was contingent on heading: odor was presented only when the fly's heading fell within a predefined angular sector relative to the visual landmark. Each session began with an initial blank period of 200~s. In Protocol A, this blank period was followed immediately by two odor blocks, yielding a three-block sequence (blank 1, odor zone 1, odor zone 2) with durations of 200, 400, and 400~s, respectively. In Protocol B, the session comprised a five-block sequence of alternating blank and odor epochs (blank 1, odor zone 1, blank 2, odor zone 2, blank 3). For Protocol B, two timing schedules were used: 200, 300, 150, 300, 150~s or 200, 350, 60, 350, 60~s, corresponding to the five successive blocks. Thus, in both protocols, odor was presented twice per session, first in one heading-defined sector and then in the opposite sector; Protocol A omitted an intervening blank period between the two odor blocks, whereas Protocol B included a brief clean-air interval between them. This design preserved a common experimental logic while varying block structure and duration to reduce stereotyped temporal expectation.

For calcium imaging experiments, trials were separated by 1~min of darkness. Because these experiments depended on spontaneously expressed behavior, trials were continued until the fly stopped walking.

\section{Definition of odor bout}

For analysis, the odor command signal was represented as a binary time series indicating valve OFF and ON states at each time point. Odor bouts were identified by first detecting individual ON epochs from OFF-to-ON and ON-to-OFF transitions in this signal. Because behaviorally contingent odor delivery often generated rapid switching of the valve during a single encounter, adjacent ON epochs separated by gaps of no more than 1~s were merged and treated as a single odor bout. Bout onset was defined as the start of the first ON epoch in the merged sequence, and bout offset as the end of the last ON epoch. For each bout, we measured both the overall bout duration, including any brief intervening OFF periods, and the cumulative odor-on duration, defined as the summed duration of all ON epochs within the bout. Bouts were included in subsequent analyses only if cumulative odor-on duration was at least 0.5~s.

\section{Computing exponentially weighted history metric}

To quantify recent odor experience at any time point during a trial, we computed an event-based exponentially weighted odor history variable. Odor encounter times were defined as transitions of the commanded valve state into the odor-on condition. For any time $t$, we identified all prior odor-on transition times $\{t_i\}$ satisfying $t_i < t$, and calculated odor history as $\sum_i \exp(-(t - t_i)/\tau)$, where $\tau$ denotes the exponential decay time constant. Under this formulation, recent odor encounters contribute more strongly than earlier encounters, with each contribution decaying continuously as a function of time since encounter onset. Because this measure is based on odor-on transition events rather than odor occupancy, it reflects the recent temporal structure and frequency of odor encounters, but does not scale with encounter duration. For bout-based analyses, this quantity was evaluated at the bout start time.

\section{Definition of sparse vs.\ dense regime}

To quantify locomotor behavior as a function of recent odor experience, we computed an event-based exponentially weighted odor-history variable continuously throughout each trial from the sequence of prior odor-on transitions using an exponential decay constant $\tau$. For each fly, the distribution of odor-history values across all trial time points was examined as a histogram, and clustered regions of this distribution were used to define odor-history regimes. The lowest-history regime, corresponding to periods before odor-history accumulation, was excluded from analysis. The highest-history regime was defined as the dense regime, reflecting periods with clustered recent odor encounters, and the intermediate regime was defined as the sparse regime, reflecting periods with intermittent but relatively infrequent recent odor encounters. All trial time points were labeled according to these regimes. This time-resolved segmentation was used to compute average locomotor statistics, including forward velocity, yaw speed, and trajectory straightness, during sparse and dense periods across the full trial. For bout-aligned analyses, each bout was assigned the label of the odor-history regime present at its start time, and these labels were used to compare locomotor response dynamics between sparse and dense odor-history conditions.

\section{Immunohistochemistry}

Brains from female flies 1--3 days posteclosion were dissected in \textit{Drosophila} external saline and fixed in 4\% paraformaldehyde (Electron Microscopy Sciences, 15714) in phosphate-buffered saline (PBS; Thermo Fisher Scientific, 46-013-CM) for 15~min at room temperature. Samples were washed in PBS and blocked for 20~min in PBS containing 5\% normal goat serum (Sigma-Aldrich, G9023) and 0.44\% Triton X-100 (Sigma-Aldrich, T8787). Brains were incubated in primary antibody diluted in blocking solution for approximately 24~h at room temperature, washed in PBS, and then incubated in secondary antibody in blocking solution for approximately 24~h at room temperature. Antibodies were selected according to the experimental protocol. After a final PBS wash, brains were mounted in antifade medium (Vectashield, Vector Laboratories, H-1000) for imaging. Images were acquired on a Leica SPE confocal microscope using a 40$\times$, 1.15~NA oil-immersion objective. Image stacks contained 50--200 z-sections collected at 1~\textmu m intervals and were recorded at a resolution of 1,024\,$\times$\,1,024 pixels. For visualization of Gal4 expression patterns, primary antibodies were chicken anti-GFP (1:1,000; Abcam, ab13970) and mouse anti-Bruchpilot (1:30; Developmental Studies Hybridoma Bank, nc82). Secondary antibodies were Alexa Fluor 488 goat anti-chicken (1:250; Invitrogen, A11039) and Alexa Fluor 633 goat anti-mouse (1:250; Invitrogen, A21050). For visualization of cell fills after whole-cell patch-clamp recordings, streptavidin-Alexa Fluor 568 (1:1,000; Invitrogen, S11226) was included in both the primary and secondary antibody solutions.

\section{Processing calcium imaging data}

Data analysis was carried out in either MATLAB 2020a or MATLAB R2023b. The calcium imaging dataset included X flies expressing GCaMP driven by MBON21-Gal4, X flies expressing GCaMP driven by MBON09-Gal4, X flies expressing GCaMP driven by the hDeltaB split-Gal4 line, X flies expressing GCaMP driven by the FB4R split-Gal4 line, X flies expressing GCaMP driven by the hDeltaB \& MBON09 split-Gal4 line, and X flies expressing GCaMP driven by the hDeltaB \& MBON21 split-Gal4 line. For each trial, rigid motion in the $x$, $y$, and $z$ dimensions was corrected using the NoRMCorre algorithm. Regions of interest (ROIs) were defined across the z-stack. For each ROI, $\Delta F/F$ was computed, with baseline fluorescence ($F$) taken as the mean of the lowest 10\% of fluorescence values within that trial, which lasted 600~s.

For imaging of the fan-shaped body, ROIs of approximately equal width were placed along the horizontal axis of the structure without overlapping. For MBON imaging, two ROIs were used: one corresponding to the left MBON and the other to the right MBON.

\section{Processing locomotion and visual arena data}

Spherical treadmill position was tracked online with machine-vision software (Fictrac v.2.1) and recorded as a voltage signal. For offline analysis, this voltage trace was converted to radians and unwrapped. The resulting signals were low-pass filtered with a second-order Butterworth filter with a 0.003 corner frequency, then downsampled to half of the ball-tracking update rate. Velocity was computed with the MATLAB \texttt{gradient} function. Spurious high velocity values exceeding 20~rad\,s$^{-1}$ were clipped to 20~rad\,s$^{-1}$, after which the timeseries were smoothed in MATLAB using the \texttt{smooth} function with the loess method and a 33~ms window, and then resampled to 60~Hz, matching the ball-tracking update rate. Forward and lateral velocities were converted to millimetres per second, while yaw velocity was expressed in degrees per second.

During calcium imaging, a signal from the imaging software marking the end of each volumetric stack was recorded on the same acquisition card as the online ball-tracking signals. The ball-kinematic data were then resampled to the imaging acquisition rate ($\sim$4--7~Hz depending on the ROI), enabling alignment with the acquired imaging volumes.
