% acknowledgments section

%\newthought{}

Many people have helped make my graduate work possible, and I’d like to thank a few in particular.

First, I want to thank Gloria Choi and Alan Jasanoff, my undergraduate advisors at MIT, for teaching me how to work in a lab and how to ask interesting questions. I also thank Jeongtae Kwon and Changyeon Ryu from the Choi lab, and He Wei from the Jasanoff lab for directly mentoring me and encouraging me to pursue research.

%From **[Institution / Research Campus Name]**, I thank **[PI / Supervisor Name]** for hiring me as a research technician when I knew very little about **[Field / Topic]**. I’m especially grateful for **[his/her/their]** patience and for giving me the space to grow in the lab. I also thank **[Mentor Name]** for helping me feel like I could make it as a scientist. I’m grateful to **[Mentor Name]**, who was first a graduate student in the **[Lab Name]** lab and later a postdoc in the **[Other Lab Name]** lab, for sharing **[his/her/their]** excitement about **[Subfield]** and for inviting me to help teach at **[Course / Program Name]** at **[Location]** during the summers. Learning by following **[him/her/them]** has been an incredibly useful tactic in my career.

At **[University Name]**, I want to thank the many wonderful people in the **[Advisor’s Lab Name]** lab. **[Person Name]**, for working with me on **[Project Description]** (which ultimately became the foundation of my thesis); **[Person Name]**, for **[quality: positivity/encouragement/etc.]**; **[Person Name]**, for providing **[calm/advice/etc.]**; **[Person Name]**, for **[quality]**; **[Person Name]**, for **[friendliness/long chats/etc.]**; **[Person Name]**, for enthusiasm for **[topic/recordings/etc.]**; **[Person Name]**, for answering questions I should have Googled; **[Person Name]**, for encouragement and camaraderie as a fellow **[shared identity/role]**; **[Person Name]**, for being a great friend and taking classes with me; **[Person Name]**, for letting me contribute to **[paper/project]**, an experience I greatly value; **[Person Name]**, for help in the lab and being a friendly colleague; **[Person Name]** (and **[Partner Name]**) for reliably making my weeks more fun and being top-notch friends; **[Person Name]**, for being a supportive friend in and out of the lab; **[Person Name]**, for **[specific small gesture that mattered]**; **[Person Name]**, for constant kindness and support; **[Person Name]**, for boundless excitement and shared nerdiness; **[Person Name]**, for bringing fun and energy to the lab; **[Person Name]**, for being an excellent colleague and deep source of knowledge; **[Person Name]**, for steady support and kindness; **[Person Name]**, for taking on the role of **[role/joke role]**; and many others, including newer lab members—especially **[Person Name]**, who has helped me in big and small ways since joining. I expect nothing but great things from these and the other talented, kind people in the lab: **[Name]**, **[Name]**, **[Name]**, **[Name]**, **[Name]**, **[Name]**, and **[Name]**.

I especially want to thank **[Former Lab Member Name]**, whose encouragement, excitement, and advice were a constant source of support throughout my graduate years, and for which I will be forever grateful.

And of course, I thank **[Advisor Name]**, my advisor, for supporting me through the extreme highs and lows of life and lab. I continue to be inspired by **[his/her/their]** enthusiasm, clarity of thought, and generosity. I’m deeply grateful to have spent my graduate career in **[his/her/their]** lab—one of the most supportive environments I have ever known—and I’m humbled to have trained under **[him/her/them]**.

From the **[Program Name]**, I want to thank **[Friend Name]** for being a close friend from interviews onward; **[Friend Name]** for making me laugh hard and often; and **[Friend Name]**, **[Friend Name]**, **[Friend Name]**, **[Friend Name]**, and **[Friend Name]** for making sure there was always something fun planned.

I also thank **[Former Administrator Name]**, the former program administrator, for **[his/her/their]** dedication to graduate students and for always being present to hear our concerns, no matter the scale.

I’m grateful to the current and former **[Program Name]** directors—**[Director Name]**, **[Director Name]**, **[Director Name]**, and **[Director Name]**—for their efforts to continuously improve the program, and I’m excited to see how future students benefit from their work.

I also thank the members of my **[Committee Name / DAC]**: **[Name]**, **[Name]**, **[Name]**, and **[Name]**. They were supportive and encouraging through what must have seemed like a long stretch of methods development.

I’m also grateful to **[Name]**, **[Name]**, **[Name]**, **[Name]**, and **[Name]** for guidance during my early **[technique]** experiments. In retrospect, I’m now certain my initial ideas must have sounded outrageous.

Finally, I thank my family. I thank my mother, **[Mother Name]**, for her unwavering commitment to my education and for supporting my choices—I’m deeply grateful for her sacrifices. I thank my **[sibling relationship]**, **[Sibling Name]**, for putting up with me and for being there through the ups and downs of graduate school. I also thank my **[grandparent relationship]**, **[Grandparent Name]**, and my **[aunt/uncle relationship]**, **[Name]** and **[Name]**, for their encouragement as I worked toward my goals.
